---
title: "marathonR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{marathonR}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
# marathonR

This packages provides methods for the uniform sampling of binary matrices with restricted row and column sums.
In particular, the package provides two functions for the uniform sampling of 
  a) binary matrices with fixed row and column sums.
  b) binary matrices whose row and column sums are restricted by lower and upper bounds.

## Fixed Margins

The function 'marathonR::SampleFixed' is designed to produce uniformly distributed binary matrices with prescribed row and column sums.
For example, the following function will produce ten random binary matrices with the specified set of row and column sums.

```{r sampleFixed}
rowsum <- c(2,2,1)
colsum <- c(1,1,1,1,1)
SampleFixed(rowsum, colsum, 10)
```

This code snippet will create independent and perfectly uniformly distributed random samples.
However, the running time of this method may grow unpleasently large when the dimension of the matrix grows. 
In such cases, we may use a different sampling method by specifying the `method` parameter and defining an associated number of `steps`.

```{r sampleFixed2}
SampleFixed(rowsum, colsum, 10, method="curveball", steps=100)
```

In this case, 100 steps of the Curveball algorithm is used to produce random samples.

## Margins in Intervals

In some situations, we may want to let the row and columns sums of a binary matrix vary in prescribed intervals.
For this purpose, the function `marathonR::SampleInterval` is designed to draw `N` independent samples from the set of binary matrices whose row and column sums lie in prescribed intervals.
For example, we might be interestend in random binary matrices whose row and column sums are allowed to exceed to ones defined above by one.
Then, we might call.

```{r sampleInterval}
rowsum.lower <- c(2,2,1)
rowsum.upper <- c(3,3,2)
colsum.lower <- c(1,1,1,1,1)
colsum.upper <- c(2,2,2,2,2)
SampleInterval(rowsum.lower, rowsum.upper, colsum.lower, colsum.upper, 10)
```

While this code snippet will create perfectly random samples, its running time may be unpleasently large when the sizes of the intervals are large.
In such cases, we may select a different sampling method, and manually provide the number of steps used to produce random samples. 

```{r sampleInterval2}
SampleInterval(rowsum.lower, rowsum.upper, colsum.lower, colsum.upper, 10, method="simple", steps=100)
```
